Title Goes Here

### 2020-02-25 paper/paper.tex
Explanation for 'delayed ERepTy', with examples including a recursive example

@@

\ttt{ERep} and \ttt{ERepTy} provide an interface for transferring values between the EDSL
world and the Haskell world:

>>>>

Only the "outermost" type will be deconstructed into these fundamental building
blocks, and further deconstruction can take place on these pieces later on. For
example, consider this type:

\begin{lstlisting}
data ComplexPair where
  ComplexPair :: Complex Double
                 -> Complex Double
                 -> ComplexPair
  deriving (Generic, Eq, Show)

instance ERep ComplexPair
\end{lstlisting}

Note that the instance definition is automatically generated from the
\ttt{Generic} instance.

Given this code, \ttt{ERep ComplexPair \typeeq (Complex Double, Complex Double)}. Here
is an example that demonstrates why this is more useful than if it *fully* deconstructed
each type into \ttt{Either}s and \ttt{(,)}:

\begin{lstlisting}
sumComplexPair ::
  ComplexPair -> Complex Double
sumComplexPair p =
  internalize (externalize
    (case p of
      ComplexPair a b -> a + b))
\end{lstlisting}

If \ttt{ERep ComplexPair} was fully deconstructed into
\ttt{((Double, Double), (Double, Double))}, we would need a \ttt{Num}
instance for \ttt{(Double, Double)}.
What we really want is to use the \ttt{Num} instance that already
exists for \ttt{Complex Double}, and this is what preserving this type
information allows us to do.

It is important to note that we are still able to further deconstruct this type
with further pattern matches, since \ttt{ERep (Complex Double) \typeeq (Double, Double)}:

\begin{lstlisting}
realSum :: ComplexPair -> Double
realSum p =
  internalize (externalize
    (case p of
      ComplexPair a b ->
        case a of
          a_real :+ _ ->
            case b of
              b_real :+ _ ->
                a_real + b_real))
\end{lstlisting}

\begin{tcolorbox}
  \textbf{TODO}: Should we use a single nested pattern match in the above block? Both
  are functional (and equivalent) from a technical perspective.
\end{tcolorbox}

An additional benefit of this is that recursive types require no special handling. For
example, consider:

\begin{lstlisting}
data IntList = Nil | Cons Int IntList
  deriving (Generic, Show)

instance ERep IntList
\end{lstlisting}

Note that \ttt{ERepTy IntList \typeeq Either () (Int, IntList)}. If \ttt{ERepTy}
attempted to "fully deconstruct" \ttt{IntList}, it would send the compiler
into an infinite loop.

This allows us to implement functions on such recursive types:

\begin{lstlisting}
isEmpty :: IntList -> Bool
isEmpty t =
  internalize (externalize
    (case t of
      Nil -> True
      Cons x xs -> False))

intListSum_helper :: (Int, IntList) -> Int
intListSum_helper p =
  internalize (externalize
    (case p of
       (acc, t) ->
         case t of
           Nil -> acc
           Cons x xs ->
             intListSum_helper (x+acc, xs)))

intListSum :: IntList -> Int
intListSum t =
  internalize (externalize
    (intListSum_helper (0, t)))
\end{lstlisting}

@@

\ttt{ERep} and \ttt{ERepTy} provide an interface for transferring values between the EDSL
expression language and the source Haskell language:
